\documentclass[12pt]{article}
\usepackage{setspace}
\singlespace
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
\usepackage{graphicx}

\title{\textbf{Comparing Vulnerabilities in Networks: 
Software Defined Networks in Personal and Enterprise Applications}}
\author{Maxwell Stratford}

\begin{document}

\maketitle

\section{Introduction}
With the ever changing landscape of the IT field, it is important to try to stay up to date on 
security vulnerabilities, and new technologies that are being developed. Something that the average
person may not be aware of is the concept of network security or vulnerabilities, even though this 
is one of the easiest ways for a hacker to gain access to devices and secure information. With the 
right setup, it is possible to have a secure network that can be used for personal or 
enterprise applications. One of the newer technologies that are being used in the field of networking
are SDN (Software Defined Network) bridges. These devices allow for a more 
flexible, efficient, and secure network. They do this through seperating the Control Plane, the decision
making portion, from the Data Plane, the portion of the network that handles the actual data transfer.
By centralizing the decision making, it abstracts the underlying hardware, and allows for more efficient
network management. The goal of this paper, is to compare the SDNs and basic network setups in terms of
enterprise and personal applications, and show how a network administrator can use SDNs to create a more 
secure network, and mitigate usual vulnerabilities. This comparison is important because it will firstly,
help to inform people about security vulnerabilities in their networks, and secondly, it will give them 
the tools to mitigate those vulnerabilities, and create a more secure network. 
Prior works to this paper, including \cite{biswas2021mitigation}, dive into enterprise applications
of the SDN, and how it can be used to mitigate large Threat Actors. This paper will expand on that, 
but will also include how to mitigate smaller scale attacks, that are easy for a Threat Actor to 
perform, but most networks are not prepared for. Towards this goal, Carnegie Mellon 
University's GHOSTS \cite{GHOSTSDocumentationGHOSTS} and VirtualBox \cite{virtualbox_docs} will
be used to simulate a network and its traffic. Then penetration testing of the networks will be 
performed to test the vulnerabilities of the network through basic attacks like DDoS (Directed
Denial of Service), MITM (Man in the Middle), and ARP (Address Resolution Protocol) Spoofing.
So in all, this paper will contribute to the fields understanding of SDNs and how they can not only
be used in enterprise applications, but also in personal network applications, as the end user is
the easiest target for a Threat Actor into any network.

\section{Background}
This section will cover the background information that is necessary to understand the rest of the paper.
This includes the different types of networking devices, software tools, network protocols, and attack types
that will be used in this paper.

\subsection{Networking Devices}
\begin{itemize}
    \item Router: A device that forwards data between different networks. For example, a home router
    will be the box from your ISP (Internet Service Provider) that connects your home network to your
    ISP's network, which will then be forwarded to the internet. 
    \item Switch: A device that acts like a router, but only forwards data within the same network. 
    Some home setups will have a small switch used to connect multiple devices to Ethernet.
    \item SDN Controller: A centralized software entity that manages network switches. It makes 
    routing decisions (control plane) and pushes rules to switches, which only forward packets 
    (data plane). This separates decision-making from packet forwarding.
    \item End Device: A device that is connected to the network that can send and/or receive data.
    This can be a computer, a phone, a printer, television, digital camera, etc.
\end{itemize}

\subsection{Software Tools}
These are the software tools that will be used in simulating the network and performing the penetration
testing. 
\begin{itemize}
    \item GHOSTS: A software tool developed by Carnegie Mellon University that allows for the 
    simulation of network traffic. It does this by using a virtual machine and an AI to emulate a user,
    for example, surfing the web. \cite{GHOSTSDocumentationGHOSTS}
    \item VirtualBox: A software tool developed by Oracle Corporation that allows for the creation of
    virtual machines. This will be used to create the network and the end devices for the simulation.
    \cite{virtualbox_docs}
    \item Packet Tracer: A software tool developed by Cisco that allows for the simulation of network 
    traffic and the creation of virtual networks. This will be used to create diagrams of the network topology.
    \cite{cisco_packet_tracer_install}
    \item Wireshark: A software tool that allows for the analysis of network traffic. \cite{wireshark_docs}
\end{itemize}

\subsection{Network Protocols}
These are some basic protocols that are used in every network, and are important to understand when 
discussing vulnerabilities in networks.
\begin{itemize}
    \item TCP/IP: Consists of two core protocols. TCP (Transmission Control Protocol) handles 
    reliable, ordered delivery of data, while IP (Internet Protocol) handles addressing and routing. 
    This suite is the foundation of internet communication, and attacks like DDoS target TCP's 
    connection-handling mechanisms.
    \item HTTP/HTTPS: Protocols used for transmitting web pages and other data between a web browser 
    and a web server.
    \item ARP: A protocol used for mapping IP addresses to MAC addresses. Because ARP lacks authentication,
     an attacker can send fake ARP replies to associate their MAC address with a legitimate IP. This is a 
     technique called ARP spoofing, used in man-in-the-middle attacks.
\end{itemize}

\subsection{Comparing Network Types}
\begin{itemize}
    \item Personal Network: A network that is used for personal use, such as a home network. These 
    networks are typically smaller and less complex than enterprise networks, and are often not as 
    secure. This leads to them being an easy target for Threat Actors.
    \item Enterprise Network: A network that is used for business purposes, such as a corporate network.
    These networks are typically larger and more complex than personal networks, and are often more
    secure. However, they are also a larger target for Threat Actors, and can be more difficult to 
    secure due to their size and complexity.
\end{itemize}

\subsection{Attack Types}
\begin{itemize}
    \item DDoS (Directed Denial of Service): An attack that floods a network with traffic in order to 
    overwhelm it and make it unavailable to legitimate users.
    \item ARP Spoofing: An attack that involves sending fake ARP replies to associate the attacker's MAC address
    with a legitimate IP address, allowing the attacker to intercept or manipulate traffic intended for that IP.
    \item MITM (Man in the Middle): An attack that involves intercepting and potentially altering communication 
    between two parties. 
\end{itemize}

Later in the paper, these two types of networks will be compared in terms of their vulnerabilities
and setup, and how SDN implementations can be used to mitigate those vulnerabilities and create 
a more secure network.

\section{Related Works}
Software Defined Networks (SDNs) are a relatively new technology, but a well researched one. There are
many papers that have been published on the topic, and many of them focus on the more intricate details
by diving into enterprise applications of the SDN. Particularly, on the topic of mitigating large scale 
attacks \cite{biswas2021mitigation}. However, prior research has seemed to overlook the home networks,
which face different threats and technical constraints. This section reviews relevant literature on 
SDN security, network vulnerabilities, and simulation tools, then identifies the gap this study addresses.

\subsection{SDN for Attack Mitigation}
Biswas \cite{biswas2021mitigation} investigated SDN based mitigation of network attacks like DDoS,
demonstrating how an SDN improved detection and response time of an attack when compared to a 
traditional network architecture. However, this study exclusively focused on enterprise networks,
leaving a gap in understanding of how SDNs can be applied to a home network, where resources are 
limited and attack vectors differ. 

\subsection{Traditional Network Vulnerabilities}
Rostami et al. \cite{InternetFacingRouterRostami} examined internet facing router services and 
found that over 40\% of known router IPv4 addresses had at least one vulnerable service exposed.
This failure to follow basic network security practices highlights the vulnerabilities of many
networks, particuarly home networks, where most users do not have the technical knowledge to 
secure their networks, or even know they need to secure their networks. This increases the need for
SDNs in home networks, as they can enforce security policies centrally. 

In summary, prior research has demonstrated SDN's effectiveness for attack mitigation in enterprise 
networks \cite{biswas2021mitigation} and documented widespread vulnerabilities in home routers 
\cite{InternetFacingRouterRostami}. However, there seems to be a gap in understanding of how SDNs 
could effect home network security. This paper aims to fill that gap by simulating both networks 
before and after the implementation of an SDN, and performing testing and penetration testing
to show the SDNs effect. The next section will show the methodology for this simulation and testing.

\section{Methodology}
To compare SDNs in enterprise and personal networks, this study uses a simulation based approach, 
where through simulating such networks with their traffic, and then performing penetration testing
on them, we can see the vulnerabilities of each network, and how the SDN can mitigate those 
vulnerabilities.

\subsection{The Simulation Environment}
VirtualBox \cite{virtualbox_docs} will be used to create virtual machines that will represent end users
and network devices. GHOSTS \cite{GHOSTSDocumentationGHOSTS} will then be installed on the simulated end
users to generate realistic network traffic. Both types of Network topologies, enterprise and personal, 
will be simulated this way, and then the SDN will be implemented in each network, and the simulation will 
be run again to see the effect of the SDN on the network's vulnerabilities.

\subsection{Attack Methodology}


\bibliographystyle{acm}
\bibliography{bibliography.bib}

\end{document}
